Semi-Parametric Manifold Clustering

### 2021-10-27 curve-fitting-methods/curve-fitting-scratch.Rmd
added headers
typo
updated writeup so far

```{r}
source('~/dev/pabm-grdpg/functions.R')
source('https://mtrosset.pages.iu.edu/Courses/675/graph.r')

laplacian.eigenmap <- function(W, d = 2, normalized = TRUE) {
  n <- nrow(W)
  if (normalized) {
    L <- normalized.laplacian(W)
  } else {
    L <- graph.laplacian(W)
  }
  L.eigen <- eigen(L, symmetric = TRUE)
  sweep(L.eigen$vectors[, seq(n - 1, n - d)], 2, 
        sqrt(L.eigen$values[seq(n - 1, n - d)]), `/`)
}

spectral.clustering <- function(W, K = 2, d = K, normalized = TRUE) {
  Y <- laplacian.eigenmap(W, d, normalized)
  kmeans(Y, K)$cluster
}
```

@@

```{r}
param.curve.quadr.2 <- function(t., 
                                beta10, beta11, beta12, 
                                beta20, beta21, beta22) {
  x1 <- beta10 + beta11 * t. + beta12 * t. ^ 2
  x2 <- beta20 + beta21 * t. + beta22 * t. ^ 2
  return(cbind(x1, x2))
}

estimate.one.t.quadr.2 <- function(x1, x2, 
                                   beta10, beta11, beta12, 
                                   beta20, beta21, beta22) {
  gamma0 <- beta10 * beta11 + beta20 * beta21 - beta11 * x1 - beta21 * x2
  gamma1 <- 2 * beta10 * beta12 + 2 * beta20 * beta22 + 
    beta11 ^ 2 + beta21 ^ 2 - 
    2 * beta12 * x1 - 2 * beta22 * x2
  gamma2 <- 3 * beta11 * beta12 + 3 * beta21 * beta22
  gamma3 <- 2 * (beta12 ^ 2 + beta22 ^ 2)
  roots <- polyroot(c(gamma0, gamma1, gamma2, gamma3))
  roots <- roots[abs(roots) == abs(Re(roots))]
  roots <- Re(roots)
  return(max(roots))
}

estimate.t.quadr.2 <- function(X, 
                               beta10, beta11, beta12, 
                               beta20, beta21, beta22) {
  apply(X, 1, function(x) {
    estimate.one.t.quadr.2(x[1], x[2], 
                           beta10, beta11, beta12, 
                           beta20, beta21, beta22)
  })
}

estimate.quadr.coefs.2 <- function(X, t.hat, intercept = TRUE) {
  sum.X <- colSums(X)
  sum.Xt <- colSums(sweep(X, 1, t.hat, `*`))
  sum.Xt2 <- colSums(sweep(X, 1, t.hat ^ 2, `*`))
  sum.t <- sum(t.hat)
  sum.t2 <- sum(t.hat ^ 2)
  sum.t3 <- sum(t.hat ^ 3)
  sum.t4 <- sum(t.hat ^ 4)

  if (intercept) {
    A.sub <- matrix(c(1, sum.t, sum.t2, 
                      sum.t, sum.t2, sum.t3, 
                      sum.t2, sum.t3, sum.t4), 
                    nrow = 3, ncol = 3)
    A <- Matrix::bdiag(A.sub, A.sub)

    b <- c(sum.X[1], sum.Xt[1], sum.Xt2[1], 
           sum.X[2], sum.Xt[2], sum.Xt2[2])
    return(solve(A, b))
  } else {
    A.sub <- matrix(c(sum.t2, sum.t3, sum.t3, sum.t4),
                    nrow = 2, ncol = 2)
    A <- Matrix::bdiag(A.sub, A.sub)
    b <- c(sum.Xt[1], sum.Xt2[1], sum.Xt[2], sum.Xt2[2])
    theta <- solve(A, b)
    return(c(0, theta[1:2], 0, theta[3:4]))
  }
}

estimate.quadr.curve.2 <- function(X, init.params, 
                                   intercept = TRUE,
                                   eps = 1e-6, maxit = 100) {
  if (missing(init.params)) {
    theta <- rnorm(6)
  } else {
    theta <- init.params
  }
  d.theta <- 1 / eps

  niter <- 0
  while (d.theta > eps) {
    theta.prev <- theta
    t.hat <- estimate.t.quadr.2(X,
                                theta[1], theta[2], theta[3],
                                theta[4], theta[5], theta[6])
    X.hat <- param.curve.quadr.2(t.hat, theta[1], theta[2], theta[3], 
                                 theta[4], theta[5], theta[6])
    theta <- estimate.quadr.coefs.2(X, t.hat, intercept)

    d.theta <- sum((theta - theta.prev) ^ 2) / sum(theta ^ 2)

    niter <- niter + 1
    if (niter >= maxit) {
      warning('failed to converge to a quadratic curve')
      break
    }
  }
  return(list(theta = theta, 
              X = X.hat,
              t = t.hat))
}
```

@@

Let $g(t) = (t^2, 2 t (1 - t) ) = (t^2, 2 t - 2 t^2)$. 
(This is the first two dimensions of the Hardy-Weinberg curve). 
Then $\theta = (0, 0, 1, 0, 2, -2)$.

@@

1. Initialize $\hat{\theta}^{(0)}$ (e.g., by drawing from a probability distribution). 
2. Estimate each $\hat{t}_i^{(s)}$ by minimizing $L(t_i, \hat{\theta}^{(s)} | x_i) = L_i = \|x_i - g(t_i | \hat{\theta}^{(s)})\|^2$.
3. Compute each $\hat{x}_i^{(s)} = g(\hat{t}_i^{(s)} | \hat{\theta}^{(s)})$
4. Estimate $\hat{\theta}^{(s+1)}$ by minimizing $L(\{\hat{t}_i^{(s)}\}, \theta | X) = \sum_i \|x_i - g(\hat{t}_i^{(s)} | \theta)\|^2$.
5. Repeat steps 2-4 until convergence.

If we restrict $g$ to be polynomials, then steps (2) and (4) have closed-form solutions.

@@

@@

In the next example, we draw $A \sim \RDPG(X)$ using the same H-W curve and sample size as above and estimate the true latent positions (up to rotation). 
In this example, we force the intercept terms to be zero.

@@

```{r, cache = TRUE}
out <- estimate.quadr.curve.2(Y,
                              init.params = c(0, 0, 1, 0, 2, -2),
                              intercept = FALSE,
                              maxit = 200)
thetahat <- out$theta
t.hat <- sort(out$t)
Yhat <- param.curve.quadr.2(t.hat, 
                            thetahat[1], thetahat[2], thetahat[3],
                            thetahat[4], thetahat[5], thetahat[6])

@@

@@

1. Assign an initial clustering (e.g., via spectral clustering).
2. Estimate the curve for each cluster.
3. Reassign the clusters by proximity to each cluster.
4. Repeat 2 and 3 until convergence.

@@

```{r}
compute.distances.quadr.2 <- function(X, theta) {
  t. <- estimate.t.quadr.2(X, 
                           theta[1], theta[2], theta[3], 
                           theta[4], theta[5], theta[6])
  X.hat <- param.curve.quadr.2(t., 
                               theta[1], theta[2], theta[3], 
                               theta[4], theta[5], theta[6])
  apply(X - X.hat, 1, function(x) sum(x ^ 2))
}

manifold.clustering.quadr.2 <- function(A, K = 2, 
                                        intercept = TRUE, 
                                        maxit = 100) {
  n <- nrow(A)

  z.hat <- spectral.clustering(A, K)
  X <- embedding(A, 2, 0)
  z.hat.prev <- sample(seq(K), n, replace = TRUE)

  niter <- 0

  while (!all(z.hat.prev == z.hat)) {
    z.hat.prev <- z.hat

    X1 <- X[z.hat == 1, ]
    X2 <- X[z.hat == 2, ]

    curve1 <- estimate.quadr.curve.2(X1, intercept = intercept)
    curve2 <- estimate.quadr.curve.2(X2, intercept = intercept)

    d1 <- compute.distances.quadr.2(X, curve1$theta)
    d2 <- compute.distances.quadr.2(X, curve2$theta)
    distances <- cbind(d1, d2)

    z.hat <- apply(distances, 1, which.min)

    niter <- niter + 1
    if (niter >= maxit) {
      warning('failed to converge to a clustering')
      break
    }
  }

  return(list(z = z.hat, 
              X = X,
              theta = list(curve1$theta, curve2$theta)))
}
```

@@

manifold.out <- manifold.clustering.quadr.2(P, 2, intercept = FALSE)
zhat.manifold <- manifold.out$z
if (mean(z == zhat.manifold) < .5) {
  zhat.manifold <- 2 - zhat.manifold
}
t. <- seq(0, 1, .01)
theta1 <- manifold.out$theta[[1]]
theta2 <- manifold.out$theta[[2]]
Y1 <- param.curve.quadr.2(t., 
                          theta1[1], theta1[2], theta1[3], 
                          theta1[4], theta1[5], theta1[6])
Y2 <- param.curve.quadr.2(t.,
                          theta2[1], theta2[2], theta2[3],
                          theta2[4], theta2[5], theta2[6])
Y <- rbind(Y1, Y2)
z. <- c(rep(1, length(t.)),
        rep(2, length(t.)))

ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], colour = factor(z))) + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  geom_line(aes(x = Y[, 1], y = Y[, 2],
                groups = factor(z.))) + 
  xlim(-1, 0) + ylim(-.5, .5) + 
  coord_fixed()
```

@@

Finally, we draw $A \sim \RDPG(X)$ from the above example and apply the clustering and curve fitting to the ASE of $A$.

@@

```{r cache = TRUE, fig.width = 4, fig.height = 4}
set.seed(12345)

manifold.out <- manifold.clustering.quadr.2(A, 2, intercept = FALSE)
zhat.manifold <- manifold.out$z
table(z, zhat.manifold)

plot(Xhat, col = z, asp = 1)
t. <- seq(-1, 1, .01)
theta1 <- manifold.out$theta[[1]]
theta2 <- manifold.out$theta[[2]]
Y1 <- param.curve.quadr.2(t., 
                          theta1[1], theta1[2], theta1[3], 
                          theta1[4], theta1[5], theta1[6])
Y2 <- param.curve.quadr.2(t.,
                          theta2[1], theta2[2], theta2[3],
                          theta2[4], theta2[5], theta2[6])
lines(Y1)
lines(Y2, col = 2)
```

>>>>

@@

```{r}
source('~/dev/pabm-grdpg/functions.R')
source('~/dev/dissertation/curve-fitting-methods/functions.R')
```

@@

Let $g(t) = (t^2, 2 t (1 - t) ) = (0 + 0 t + t^2, 0 + 2 t - 2 t^2)$. 
(This is the first two dimensions of the Hardy-Weinberg curve). 
Then $\theta = (0, 0, 1, 0, 2, -2)$.

@@

1. Initialize $\hat{\theta}^{(0)}$ (e.g., randomly). 
2. Estimate each $\hat{t}_i^{(s)}$ by minimizing $L(t_i, \hat{\theta}^{(s)} | x_i) = L_i = \|x_i - g(t_i | \hat{\theta}^{(s)})\|^2$.
3. Compute each $\hat{x}_i^{(s)} = g(\hat{t}_i^{(s)} | \hat{\theta}^{(s)})$
4. Estimate $\hat{\theta}^{(s+1)}$ by minimizing $L(\{\hat{t}_i^{(s)}\}, \theta | X) = \sum_i \|x_i - g(\hat{t}_i^{(s)} | \theta)\|^2$.
5. Repeat steps 2-4 until convergence.

If we restrict $g$ to be polynomials, then steps (2) and (4) have closed-form solutions. 
Alternatively, we can estimate $g$ using more general forms, e.g., splines, which may require approximation.

@@

# Estimation with Noise

@@

In the next example, we draw $A \sim \RDPG(X)$ using the same H-W curve and sample size as above and estimate the true latent positions (up to rotation).

@@

```{r, cache = TRUE}
out1 <- estimate.quadr.curve.2(Y,
                              init.params = c(0, 0, 1, 0, 2, -2),
                              intercept = TRUE,
                              maxit = 200)
thetahat <- out1$theta
t.hat <- sort(out1$t)
Yhat <- param.curve.quadr.2(t.hat, 
                            thetahat[1], thetahat[2], thetahat[3],
                            thetahat[4], thetahat[5], thetahat[6])

@@

# Clustering

@@

1. Assign an initial clustering (e.g., via spectral clustering).
2. Estimate the curve for each cluster.
3. Reassign the clusters by proximity to each curve.
4. Repeat 2 and 3 until convergence.

@@

@@

manifold.out <- manifold.clustering.quadr.2(P, 2, 
                                            intercept = FALSE, 
                                            method = 'polynomial')
zhat.manifold <- manifold.out$z
if (mean(z == zhat.manifold) < .5) {
  zhat.manifold <- 2 - zhat.manifold
}
t. <- seq(0, 1, .01)
theta1 <- manifold.out$theta[[1]]
theta2 <- manifold.out$theta[[2]]
Y1 <- param.curve.quadr.2(t., 
                          theta1[1], theta1[2], theta1[3], 
                          theta1[4], theta1[5], theta1[6])
Y2 <- param.curve.quadr.2(t.,
                          theta2[1], theta2[2], theta2[3],
                          theta2[4], theta2[5], theta2[6])
Y <- rbind(Y1, Y2)
z. <- c(rep(1, length(t.)),
        rep(2, length(t.)))

ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], colour = factor(z))) + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  # geom_line(aes(x = Y[, 1], y = Y[, 2],
  #               groups = factor(z.))) + 
  xlim(-1, 0) + ylim(-.5, .5) + 
  coord_fixed()
```

Next, we draw $A \sim \RDPG(X)$ from the above example and apply the clustering and curve fitting to the ASE of $A$.
The community detection error in this example is 30\%.
In the following plot, the points are colored according to their true labels.

```{r cache = TRUE, fig.width = 3, fig.height = 3}
set.seed(12345)

manifold.out <- manifold.clustering.quadr.2(A, 2, 
                                            intercept = FALSE,
                                            method = 'polynomial')
zhat.manifold <- manifold.out$z
# table(z, zhat.manifold)

# plot(Xhat, col = z, asp = 1)
t. <- seq(-1, 1, .01)
theta1 <- manifold.out$theta[[1]]
theta2 <- manifold.out$theta[[2]]
Y1 <- param.curve.quadr.2(t., 
                          theta1[1], theta1[2], theta1[3], 
                          theta1[4], theta1[5], theta1[6])
Y2 <- param.curve.quadr.2(t.,
                          theta2[1], theta2[2], theta2[3],
                          theta2[4], theta2[5], theta2[6])
Y <- rbind(Y1, Y2)
z. <- c(rep(1, length(t.)),
        rep(2, length(t.)))
# lines(Y1)
# lines(Y2, col = 2)

ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], colour = factor(z))) + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  geom_path(aes(x = Y[, 1], y = Y[, 2],
                groups = factor(z.))) + 
  xlim(-.25, 1.25) + ylim(-.75, .75) + 
  coord_fixed()
```

A possibly more robust modification to this is to use Bezier curves for $g$. 
This is the same functional form as the polynomial curves we used before, but with different basis functions. 
For the following examples, we consider the quadratic Bezier curve, which has the following basis form:

$$
g(t) = (1 - t)^2 p_0 + 2 t (1 - t) p_1 + t^2 p_2
$$

Where as before, $g : [0, 1] \mapsto \mathbb{R}^d$ and $t \in [0, 1]$, and each $p_r \in \mathbb{R}^d$. Thus if we can fit each $p_r$, then the procedure is the same as before.

Common methods for Bezier curve fitting do not require specification of $t_1, ..., t_n$. However, an ordering is required. The curve fitting method is then modified as follows:

1. Initialize $\hat{\theta}^{(0)}$ (e.g., randomly). 
2. Estimate each $\hat{t}_i^{(s)}$ by minimizing $L(t_i, \hat{\theta}^{(s)} | x_i) = L_i = \|x_i - g(t_i | \hat{\theta}^{(s)})\|^2$.
3. Reorder $X$ according to $t_1, ..., t_n$.
4. Estimate $\hat{p}_0^{(s)}, \hat{p}_1^{(s)}, \hat{p}_2^{(s)}$ for the reordered $X$ using Bezier curve fitting.
5. Compute $\hat{\theta}^{(s+1)}$ from $\{\hat{p}_r^{(s)}\}$.
6. Repeat steps 2-5 until convergence.

@@

```{r}
f1 <- function(t) {
  x1 <- t ^ 2
  x2 <- 2 * t - 2 * t ^ 2
  return(cbind(x1, x2))
}

f2 <- function(t) {
  x1 <- 2 * t - 2 * t ^ 2
  x2 <- 1 - 2 * t + t ^ 2
  return(cbind(x1, x2))
}
```

We have two intersecting curves, $g_1(t) = \begin{bmatrix} t^2 & 2 t (1 - t) \end{bmatrix}^\top$ and $g_2(t) = \begin{bmatrix} 2 t (1 - t) & (1 - t) ^ 2 \end{bmatrix}^\top$. $n_1 = n_2 = 128$ points are drawn uniformly from each.

```{r}
n1 <- 2 ** 7
n2 <- n1
n <- n1 + n2
z <- c(rep(1, n1), rep(2, n2))

a <- 1
b <- 1
t1 <- rbeta(n1, a, b)
t2 <- rbeta(n2, a, b)
T <- rbind(t1, t2)

X1 <- f1(t1)
X2 <- f2(t2)
X <- rbind(X1, X2)

ggplot() + 
  geom_point(aes(x = X[, 1], y = X[, 2], colour = factor(z))) + 
  coord_fixed() + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw()
```

We draw $A \sim \RDPG(X)$ and obtain the following ASE:

```{r}
P <- X %*% t(X)
diag(P) <- 0
A <- draw.graph(P)
Xhat <- embedding(A, 2, 0)
if (mean(Xhat[, 1]) < 0) Xhat[, 1] <- -Xhat[, 1]

ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], colour = factor(z))) + 
  coord_fixed() + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw()
```

Fitting two quadratic Bezier curves to these data yields a community detection error rate of 10\%. 
In the following plot, the points are labeled according to their estimated labels.

```{r cache = TRUE}
set.seed(314159265)
manifold.out <- manifold.clustering.quadr.2(A, 2, 
                                            # initialization = z,
                                            # initialization = 'spectral',
                                            initialization = 'random', 
                                            method = 'bezier')
zhat.manifold <- manifold.out$z

t. <- seq(-1, 1, .01)
theta1 <- manifold.out$theta[[1]]
theta2 <- manifold.out$theta[[2]]
Y1 <- param.curve.quadr.2(t., 
                          theta1[1], theta1[2], theta1[3], 
                          theta1[4], theta1[5], theta1[6])
Y2 <- param.curve.quadr.2(t.,
                          theta2[1], theta2[2], theta2[3],
                          theta2[4], theta2[5], theta2[6])
Y <- rbind(Y1, Y2)
z. <- c(rep(1, length(t.)),
        rep(2, length(t.)))

ggplot() + 
  geom_point(aes(x = Xhat[, 1], y = Xhat[, 2], colour = factor(zhat.manifold))) + 
  labs(x = expression(x[1]), y = expression(x[2]), colour = NULL) + 
  theme_bw() + 
  geom_path(aes(x = Y[, 1], y = Y[, 2],
                groups = factor(z.))) + 
  xlim(-.25, 1.25) + ylim(-.75, .75) + 
  coord_fixed()
```

@@

### 2021-10-27 curve-fitting-methods/curve-fitting-scratch.tex
added headers
typo
updated writeup so far

Let $g(t) = (t^2, 2 t (1 - t) ) = (t^2, 2 t - 2 t^2)$. 
(This is the first two dimensions of the Hardy-Weinberg curve). 
Then $\theta = (0, 0, 1, 0, 2, -2)$.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Initialize \(\hat{\theta}^{(0)}\) (e.g., by drawing from a probability
  distribution).
\item
  Estimate each \(\hat{t}_i^{(s)}\) by minimizing
  \(L(t_i, \hat{\theta}^{(s)} | x_i) = L_i = \|x_i - g(t_i | \hat{\theta}^{(s)})\|^2\).
\item
  Compute each
  \(\hat{x}_i^{(s)} = g(\hat{t}_i^{(s)} | \hat{\theta}^{(s)})\)
\item
  Estimate \(\hat{\theta}^{(s+1)}\) by minimizing
  \(L(\{\hat{t}_i^{(s)}\}, \theta | X) = \sum_i \|x_i - g(\hat{t}_i^{(s)} | \theta)\|^2\).
\item
  Repeat steps 2-4 until convergence.
\end{enumerate}

If we restrict \(g\) to be polynomials, then steps (2) and (4) have
closed-form solutions.

@@

\begin{center}\includegraphics{curve-fitting-scratch_files/figure-latex/unnamed-chunk-4-1} \end{center}

@@

@@

In the next example, we draw $A \sim \mathrm{RDPG}(X)$ using the same H-W curve and sample size as above and estimate the true latent positions (up to rotation). 
In this example, we force the intercept terms to be zero.

@@

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Assign an initial clustering (e.g., via spectral clustering).
\item
  Estimate the curve for each cluster.
\item
  Reassign the clusters by proximity to each cluster.
\item
  Repeat 2 and 3 until convergence.
\end{enumerate}

@@

@@

\begin{center}\includegraphics{curve-fitting-scratch_files/figure-latex/unnamed-chunk-8-1} \end{center}

@@

Finally, we draw $A \sim \mathrm{RDPG}(X)$ from the above example and apply the clustering and curve fitting to the ASE of $A$.

@@

\begin{verbatim}
   zhat.manifold
z     1   2
  1 187  69
  2  87 169
\end{verbatim}

>>>>

Let $g(t) = (t^2, 2 t (1 - t) ) = (0 + 0 t + t^2, 0 + 2 t - 2 t^2)$. 
(This is the first two dimensions of the Hardy-Weinberg curve). 
Then $\theta = (0, 0, 1, 0, 2, -2)$.

\begin{center}\includegraphics{curve-fitting-scratch_files/figure-latex/unnamed-chunk-2-1} \end{center}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Initialize \(\hat{\theta}^{(0)}\) (e.g., randomly).
\item
  Estimate each \(\hat{t}_i^{(s)}\) by minimizing
  \(L(t_i, \hat{\theta}^{(s)} | x_i) = L_i = \|x_i - g(t_i | \hat{\theta}^{(s)})\|^2\).
\item
  Compute each
  \(\hat{x}_i^{(s)} = g(\hat{t}_i^{(s)} | \hat{\theta}^{(s)})\)
\item
  Estimate \(\hat{\theta}^{(s+1)}\) by minimizing
  \(L(\{\hat{t}_i^{(s)}\}, \theta | X) = \sum_i \|x_i - g(\hat{t}_i^{(s)} | \theta)\|^2\).
\item
  Repeat steps 2-4 until convergence.
\end{enumerate}

If we restrict \(g\) to be polynomials, then steps (2) and (4) have
closed-form solutions. Alternatively, we can estimate \(g\) using more
general forms, e.g., splines, which may require approximation.

@@

@@

\hypertarget{estimation-with-noise}{%
\section{Estimation with Noise}\label{estimation-with-noise}}

@@

In the next example, we draw $A \sim \mathrm{RDPG}(X)$ using the same H-W curve and sample size as above and estimate the true latent positions (up to rotation).

@@

\begin{center}\includegraphics{curve-fitting-scratch_files/figure-latex/unnamed-chunk-5-1} \end{center}

\hypertarget{clustering}{%
\section{Clustering}\label{clustering}}

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Assign an initial clustering (e.g., via spectral clustering).
\item
  Estimate the curve for each cluster.
\item
  Reassign the clusters by proximity to each curve.
\item
  Repeat 2 and 3 until convergence.
\end{enumerate}

@@

Next, we draw $A \sim \mathrm{RDPG}(X)$ from the above example and apply the clustering and curve fitting to the ASE of $A$.
The community detection error in this example is 30\%.
In the following plot, the points are colored according to their true labels.

\begin{center}\includegraphics{curve-fitting-scratch_files/figure-latex/unnamed-chunk-7-1} \end{center}

@@

A possibly more robust modification to this is to use Bezier curves for
\(g\). This is the same functional form as the polynomial curves we used
before, but with different basis functions. For the following examples,
we consider the quadratic Bezier curve, which has the following basis
form:

\[
g(t) = (1 - t)^2 p_0 + 2 t (1 - t) p_1 + t^2 p_2
\]

Where as before, \(g : [0, 1] \mapsto \mathbb{R}^d\) and
\(t \in [0, 1]\), and each \(p_r \in \mathbb{R}^d\). Thus if we can fit
each \(p_r\), then the procedure is the same as before.

Common methods for Bezier curve fitting do not require specification of
\(t_1, ..., t_n\). However, an ordering is required. The curve fitting
method is then modified as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  Initialize \(\hat{\theta}^{(0)}\) (e.g., randomly).
\item
  Estimate each \(\hat{t}_i^{(s)}\) by minimizing
  \(L(t_i, \hat{\theta}^{(s)} | x_i) = L_i = \|x_i - g(t_i | \hat{\theta}^{(s)})\|^2\).
\item
  Reorder \(X\) according to \(t_1, ..., t_n\).
\item
  Estimate \(\hat{p}_0^{(s)}, \hat{p}_1^{(s)}, \hat{p}_2^{(s)}\) for the
  reordered \(X\) using Bezier curve fitting.
\item
  Compute \(\hat{\theta}^{(s+1)}\) from \(\{\hat{p}_r^{(s)}\}\).
\item
  Repeat steps 2-5 until convergence.
\end{enumerate}

@@

We have two intersecting curves, $g_1(t) = \begin{bmatrix} t^2 & 2 t (1 - t) \end{bmatrix}^\top$ and $g_2(t) = \begin{bmatrix} 2 t (1 - t) & (1 - t) ^ 2 \end{bmatrix}^\top$. $n_1 = n_2 = 128$ points are drawn uniformly from each.

We draw $A \sim \mathrm{RDPG}(X)$ and obtain the following ASE:

\begin{center}\includegraphics{curve-fitting-scratch_files/figure-latex/unnamed-chunk-10-1} \end{center}

Fitting two quadratic Bezier curves to these data yields a community detection error rate of 10\%. 
In the following plot, the points are labeled according to their estimated labels.

\begin{center}\includegraphics{curve-fitting-scratch_files/figure-latex/unnamed-chunk-11-1} \end{center}

@@
